---
title: "apply-prep"
author: "Azul Carrillo"
date: "2022-11-15"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reusing code - Apply function - The apply family of functions

Introduction to repeating things in R.
```{r}
est_mass <- function(volume) { 
  mass <- 2.65 * volume^0.9
  return(mass)
}

est_mass(1.6)
est_mass(5.6)


```

Using vectorized functions.
```{r}
c(1, 2, 3) * 2 #element wise calc 
volumes <- c( 1.6, 5.6, 3.1)
est_mass(volumes)

library(stringr)

str_to_sentence(c("dipodomys", "chaetodipus")) # capitalazes the first letter of a sentence 
genus <- c("dipodomys", "chaetodipus", "dipodomys")
species <- c("ordii","baileyi","spectabilis")

combined_genus_species <- function(genus, species){
  genus_cap <-  str_to_sentence(genus)
  genus_species <- paste(genus_cap, species)
  return(genus_species)
}
combined_genus_species(genus, species)
data <- data.frame(genus, species)  
combined_genus_species(data$genus, data$species)
```

Apply functions
```{r eval = FALSE}
#apply the function to each item in the vector and return a vector or list of th e same size this doesnt recure that the calc work on vectors 
est_mass <- function(volume) { 
  if(volume > 5 ){
   mass <- 2.65 * volume^0.9 
  } else { mass <- NA
    }
  return(mass)
}
volumes <- c( 1.6, 5.6, 3.1)
est_mass(volumes)
sapply(volumes, est_mass) # first argument is a single vector that has the volume and the second argument is the function we want to apply to the vector 

#what the above sapply is doing under the surface 
c(est_mass(volumes[1]), est_mass(volumes[2]), est_mass(volumes[3]))
```

The mapply function For Functions with Multiple Vector Arguments
```{r}
est_mass <- function(volume, veg_type) { 
  if(veg_type == "tree" ){
   mass <- 2.65 * volume^0.9 
  } else { mass <- NA
    }
  return(mass)
}
volumes <- c( 1.6, 5.6, 3.1)
veg_type <- c("shrub", "tree", "tree")
#function is the firat argument, the name of the argument = the name our vectoror object in the outter program , for veg type theyre both name hte same 
mapply(FUN = est_mass, volume = volumes, veg_type = veg_type)
mapply(est_mass, volumes, veg_type)
```
 
Combining functions with dplyr - every row in a data frame  or for every group by using group by 

to repeat things once for esch row to create new columns we use mutate 
if func is not vectorized we need to add rowwide 
we can repeat things by using group by 
```{r}
library(dplyr)
est_mass <- function(volume, veg_type) { 
  if(veg_type == "tree" ){
   mass <- 2.65 * volume^0.9 
  } else { mass <- NA
    }
  return(mass)
}
volumes <- c( 1.6, 5.6, 3.1)
veg_type <- c("shrub", "tree", "tree")
plant_data <- data.frame (volumes, veg_type)

est_mass_vectorized <- function(volume){
  mass <- 2.65 * volume^0.9 
  return(mass)
}

plant_data %>%
  mutate(masses = est_mass_vectorized(volumes))

#working with non vectorized functions this wont work unless uyou add rowwise function 
plant_data %>%
  rowwise() %>%
  mutate(masses = est_mass(volumes, veg_type))


get_biomass <-  function(volumes) {
  masses <- est_mass_vectorized(volumes)
  biomass <- sum(masses)
  return(biomass)
}
get_biomass(volumes)

plant_data%>%
  group_by(veg_type)%>%
  summarize(biomass= get_biomass(volumes))

```

```{r}
```