---
title: "apply functions"
author: "Azul Carrillo"
date: "2022-11-15"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Write a function named mass_from_length_theropoda() that takes length as an argument to get an estimate of mass for Theropoda dinosaurs. Use the equation mass <- 0.73 * length^3.63. Copy and run the code below to generate the object theropoda_lengths in your R environment. Pass the entire vector to your function (by giving it as value for the length argument); this calculates the mass for each length value in the vector theropoda_lengths.
```{r}
mass_length_theropoda <- function(length){
   mass <- 0.73 * length^3.63
   return(mass)
}
mass_length_theropoda(100) #13283816
```

```{r}
theropoda_lengths <- c(17.8013631070471, 20.3764452071665, 14.0743486294308, 25.65782386974, 26.0952008049675, 20.3111541103134, 17.5663244372533, 11.2563431277577, 20.081903202614, 18.6071626441984, 18.0991894513166, 23.0659685685892, 20.5798853467837, 25.6179254233558, 24.3714331573996, 26.2847248252537, 25.4753783544473, 20.4642089867304, 16.0738256364701, 20.3494171706583, 19.854399305869, 17.7889814608919, 14.8016421998303, 19.6840911485379, 19.4685885050906, 24.4807784966691, 13.3359960054899, 21.5065994598917, 18.4640304608411, 19.5861532398676, 27.084751999756, 18.9609366301798, 22.4829168046521, 11.7325716149514, 18.3758846100456, 15.537504851634, 13.4848751773738, 7.68561192214935, 25.5963348603783, 16.588285389794)
```

```{r}
sapply( theropoda_lengths, mass_length_theropoda)
```

```{r}
mass_from_length <-  function(length, a = 0.73, b=3.63 ){
  mass <- a * length^b
   return(mass)
}

sapply( theropoda_lengths, mass_from_length) 

```

Create a new version of the function named mass_from_length() that uses the equation mass <- a * length^b and takes length, a and b as arguments. In the function arguments, set the default values for a to 0.73 and b to 3.63. If you run this function with just the length data from Part 1, you should get the same result as Part 1. Copy the data below into R and call your function using the vector of lengths from Part 1 (above) and these vectors of a and b values to estimate the mass for the dinosaurs using different values of a and b.

```{r}
a_values <- c(0.759, 0.751, 0.74, 0.746, 0.759, 0.751, 0.749, 0.751, 0.738, 0.768, 0.736, 0.749, 0.746, 0.744, 0.749, 0.751, 0.744, 0.754, 0.774, 0.751, 0.763, 0.749, 0.741, 0.754, 0.746, 0.755, 0.764, 0.758, 0.76, 0.748, 0.745, 0.756, 0.739, 0.733, 0.757, 0.747, 0.741, 0.752, 0.752, 0.748)

```

```{r}
b_values <- c(3.627, 3.633, 3.626, 3.633, 3.627, 3.629, 3.632, 3.628, 3.633, 3.627, 3.621, 3.63, 3.631, 3.632, 3.628, 3.626, 3.639, 3.626, 3.635, 3.629, 3.642, 3.632, 3.633, 3.629, 3.62, 3.619, 3.638, 3.627, 3.621, 3.628, 3.628, 3.635, 3.624, 3.621, 3.621, 3.632, 3.627, 3.624, 3.634, 3.621)
```

```{r}
mapply(mass_from_length, theropoda_lengths, a_values, b_values)

mass_from_length(length = theropoda_lengths, a = a_values, b = b_values)
```
creating a data frame 
```{r}
library(dplyr)
dino_data <- data.frame(theropoda_lengths, a_values, b_values)

dino_data %>%
  mutate(masses = mass_from_length(theropoda_lengths, a_values, b_values))

```

```{r}
mass_from_length_max <- function(length){
  if(length < 20 ){
  mass <- 0.73 * length^3.63
  } else { mass <- NA
    }
  return(mass)
}
```

```{r}
sapply(theropoda_lengths, mass_from_length_max)
```

```{r}
dinosaur_lengths <- read.csv("../raw-data/dinosaur_lengths.csv")
#dinosaur_lengths
```

```{r}
get_mass_from_length_by_name <- function(length, name){
  mass <- a * length^b
   if (name == "Stegosauria") {
     a <- 10.95 
     b <- 2.64 
   } else if (name == "Theropoda") {
     a <- 0.73
     b <- 3.63
   } else if (name == "Sauropoda") {
     a <- 214.44
     b <- 1.46
 } else { 
   return(NA)
  }
     return(mass)
}




```

```{r}
get_mass_from_length_by_name <- function(length, name){
   if (name == "Stegosauria") {
     a <- 10.95 
     b <- 2.64 
   } else if (name == "Theropoda") {
     a <- 0.73
     b <- 3.63
   } else if (name == "Sauropoda") {
     a <- 214.44
     b <- 1.46
 } else {
   a <- NA
   b <- NA
  }
     mass <- a * length^b
     return(mass)
}


```

```{r}
mapply(get_mass_from_length_by_name, dinosaur_lengths$lengths, dinosaur_lengths$species)

```
Use this function and mapply() to calculate the estimated mass for each dinosaur. You’ll need to pass the data to mapply() as single vectors or columns, not the whole data frame.

Using dplyr, add a new masses column to the data frame (using rowwise(), mutate() and your function) and print the result to the console.
Using ggplot2, make a histogram of dinosaur masses with one subplot for each species (remember facet_wrap()).
```{r}
library(dplyr)
library(ggplot2)
dinosaur_lengths %>%
  rowwise() %>%
  mutate(masses = get_mass_from_length_by_name(lengths, species)) %>%
  ggplot(mapping = aes(x = masses)) +
  geom_histogram() +
  facet_wrap(~species)
```
```{r}

```