---
title: ' Making Choices'
author: "Azul Carrillo"
date: "2022-11-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Logical and conditional statements 
both of there return TRUE or FALSE values 
some examples of logical statements 

```{r}
1 == 1 
1 > 1
1 < 1
1 >= 1
1 <= 1

"a"=="a"
"a"<"a"
"a">"a"
"apple" < "banana"

c(1, 2, 3, 4, 1, 1) == 1 


letters < "m"
letters[letters < "m"]
```
load data from the uhuru dataset
```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
species <- read.csv(file = "../raw-data/species.csv")

colnames(surveys)

```

```{r}
library(dplyr)
combined <- dplyr::inner_join(surveys, species, by = "species_id")
```

```{r, eval = FALSE}
#combined$year
#combined[,4]
#combined[]

combined$year == 1977
combined$genus == "Dipodomys"

data.frame(combined$genus, combined$genus == "Dipodomys")
data.frame(combined$genus, combined$genus == "Dipodomys")
```
a vector that test species to 
```{r, eval = FALSE}
combined$species == "merriami"
data.frame(combined$species, combined$species == "merriami")
```
examples of conditional statements
uses the symbols & , |
```{r}
1 == 1 & 1 < 2
1 == 1 & 1 == 2
1 == 1 | 1 == 2
```

```{r}
#combined$year == 1977 & combined$genus == "Dipodomys" & combined$species == "merriami"
```
#1 
```{r}
w <- 10.2
x <- 1.3
y <- 2.8
z <- 17.5
colors <- c("red", "blue", "green")
masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
dna1 <- "attattaggaccaca"
dna2 <- "attattaggaacaca"


w > 10 # TRUE
colors =='green' #FALSE FALSE  TRUE short answer is yes green is in the 3rd place 
x > y # FALSE
masses > 40 # TRUE FALSE FALSE  TRUE  TRUE no only 3/5 are greater than 40
2 * x + 0.2 == y #FALSE
dna1 == dna2#FALSE
dna1 != dna2# TRUE
w > x | y > z # TRUE
x* w #is between 13.2 and 13.5 TRUE [1] 13.26
masses >= 30 & masses <= 50 #TRUE  TRUE FALSE FALSE  TRUE
```
#2

```{r}

age_class = "sapling"
 if (age_class == "sapling"){
   y = 10
 }
 y
```
#3
```{r}
age_class <- "seedling"
if(age_class == "sapling"){
  y <- 10 
} else{
  y <- 5
}

y 
```
#4
```{r}
age_class = "adult"
if (age_class == "sapling"){
   y <- 10 
 } else if  (age_class == "seedling"){
   y <-  5 
   }else { 
 y <-  0 
}
 y
```
#5 
```{r}
get_y = function(age_class){
  if (age_class == "sapling"){
   y <- 10 
 } else if  (age_class == "seedling"){
   y <-  5 
   }else { 
 y <-  0 
   }
  return(y)
}
 
get_y("sapling") #10 
get_y("seedling") # 5
get_y("adult") # 0 
get_y ("mature") #0 
get_y("established")#0 

```
#6 


```{r}
get_mass_from_length <- function(length, a = 39.9, b = 2.6 ){
  mass <- a * length ^ b
  return(mass)
}


```

Create a new function get_mass_from_length_by_name() that:
takes two arguments, the length and the name of the dinosaur group.
Inside this function use if/else if/else statements to check if the name is one of the following values and if so set a and b to the appropriate values (from Seebacher 2001).
Stegosauria: a = 10.95 and b = 2.64.
Theropoda: a = 0.73 and b = 3.63.
Sauropoda: a = 214.44 and b = 1.46.
If the name is not any of these values, the function should return NA. - Once the function has chosen the correct a and b values, have it run get_mass_from_length() with the appropriate values and return the estimated mass.

```{r}
#6.1
get_mass_from_length_by_name = function(length,name) {
   if (name == "Stegosauria") {
     a <- 10.95
     b <- 2.64 
   } else if (name == "Theropoda") {
     a <- 0.73
     b <- 3.63
   } else if (name == "Sauropoda") {
     a <- 214.44
     b <- 1.46
   } else { 
   return(NA)
   }
 mass<- get_mass_from_length(length = length, a = a, b = b )
 return(mass)
}  
#6.2
get_mass_from_length_by_name(10, "Stegosauria")#4779.848
get_mass_from_length_by_name(8, "Theropoda")#1385.286
get_mass_from_length_by_name(12, "Sauropoda")#8070.685
get_mass_from_length_by_name(13, "Ankylosauria") #NA 
```
#6.3 challenge 
```{r}
get_mass_from_length_by_name = function(length,name) {
   if (name == "Stegosauria") {
     a <- 10.95
     b <- 2.64 
   } else if (name == "Theropoda") {
     a <- 0.73
     b <- 3.63
   } else if (name == "Sauropoda") {
     a <- 214.44
     b <- 1.46
   } else { 
   message("No known estimation for ", name)
   return(NA)
   }
 mass<- get_mass_from_length(length = length, a = a, b = b )
 return(mass)
}  
get_mass_from_length_by_name(13, "Ankylosauria") #No known estimation for Ankylosauria [1] NA
```

#6.4 challenge 
```{r}
get_mass_from_length_by_name = function(length,name) {
   if (name == "Stegosauria"& length > 8) {
     a <- 10.95 
     b <- 2.64 
    } else if (name == "Stegosauria" & length < 8) {
     a <- 8.5
     b <- 2.8
   } else if (name == "Theropoda") {
     a <- 0.73
     b <- 3.63
   } else if (name == "Sauropoda") {
     a <- 214.44
     b <- 1.46
   } else { 
   message("No known estimation for ", name)
   return(NA)
   }
 mass<- get_mass_from_length(length = length, a = a, b = b )
 return(mass)
}  
get_mass_from_length_by_name(6, "Stegosauria")# 1283.047
```
