---
title: "Creating functions"
author: "Azul Carrillo"
date: "2022-11-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Intro to creating functions 
Why is it good to create functions - 
create more columns in a data frame with a specfic req
creating functions can help you understand and be able to make changes in order to obtain the desired output stay organized 
copy and pasting is very error prone 

```{r}
convert_pounds_to_grams <- function(pounds) {
    grams = 453.6 * pounds
    return(grams)
}
```

```{r}
convert_pounds_to_grams <- function(pounds) {
   grams= 453.6 * pounds
    return(grams)
}
#convert_pounds_to_grams() #says value is missing 
convert_pounds_to_grams(3.75)
convert_pounds_to_grams(pounds = 3.75)


```

```{r}
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}
get_mass_from_length_theropoda(16) #17150.56
```

```{r}
get_mass_from_length <- function(length){
  mass <- 214.44 * length ^ 1.46
  return(mass)
}
get_mass_from_length(26)#24955.54kg
```

```{r}
get_mass_from_length <- function(length, a = 39.9, b = 2.6 ){
  mass <- a * length ^ b
  return(mass)
}
get_mass_from_length(length=22,a = 214.44, b = 1.46)#19554.33
```

```{r}
get_mass_from_length <- function(length, a = 39.9, b = 2.6 ){
  mass <- a * length ^ b
  return(mass)
}
get_mass_from_length(length=16) # 53911.93
```
Write a function called convert_kg_to_pounds that converts kilograms into pounds (pounds = 2.205 * kg).
Use that function and your get_mass_from_length() function from Default Arguments to estimate the weight, in pounds, of a 12 m long Stegosaurus with a = 10.95 and b = 2.64 (The estimated a and b values for Stegosauria from Seebacher 2001).
```{r}
get_mass_from_length <- function(length, a = 39.9, b = 2.6 ){
  mass <- a * length ^ b
  return(mass)
}

convert_kg_to_pounds <- function(kg){
 pounds <- 2.205 * kg
  return(pounds)
}

library(dplyr)
weight_in_lbs <- get_mass_from_length(12, 10.95, 2.64) %>%
  convert_kg_to_pounds() #17055.37
weight_in_lbs

convert_kg_to_pounds(5)
```

```{r}

```