---
title: "functions-after"
author: "Azul Carrillo"
date: "2022-11-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# A function for the Portal data set

1. loading the needed data into r with the read.csv func and asigning it a name 

```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
species <- read.csv(file = "../raw-data/species.csv")
plots <- read.csv(file = "../raw-data/plots.csv") 
```
2. first loading that package that has the inner join function so that it can be used combined the survey and species tables into a single data frame using the inner_join function joined them by species 
```{r}
library(dplyr)
surveys_species <- inner_join(surveys, species, by = "species_id")
```

need the col names in order to add the correct argument 
```{r}
colnames(surveys_species)

```
3. Write a function that:
Takes three arguments - a data frame (the combined table created before), a genus name, and a species name,
Uses functions from the dplyr package to produce a data frame with two columns: year and count, where count is the number of individuals (i.e., the number of rows) for the species indicated by genus and species in that year,
Returns the resulting data frame. Note that this data frame is a time series by year.

gives info for all genus adn all years but addind the filter it changes it 
```{r}
library(dplyr)
func_1 <- function(surveys_species, genus , species){
results <- surveys_species %>%
  group_by(year, genus, species) %>%
  summarise(count = n())
  return(results[ ,c("year", "count")])
}

results <- surveys_species %>%
  group_by(year, genus, species) %>%
  summarise(count = n())
  

results[ ,c("year", "count")]
```
#try 2 
added the filter to the function to specify what we want from the genus column and the species column 
removed the subsetting at the return function because it was cousing an issue in the output gruped only by year using the group_by 
```{r}
library(dplyr)
func_2 <- function(combined, genus_name , species_name){
results <- combined %>%
  filter(genus == genus_name , species == species_name) %>%
  group_by(year) %>%
  summarise(count = n())
  return(results)
}

head(surveys_species)
```

running a pipe befor creating the function to see that our function is doing the correct command 
```{r}
results <- surveys_species %>%
  filter(genus == "Dipodomys", species == "merriami" ) %>%
  group_by(year) %>%
  summarise(count = n())


```
#4
creating a new object that uses our function to create a time series for  the genus "Dipodomys" and species "merriami"
```{r}
plot1 <- func_2(surveys_species, "Dipodomys", "merriami")
plot1
```
#5
creating a new object that uses our function to create a time series for  the genus "Chaetodipus" species "penicillatus"
```{r}
plot2 <- func_2(surveys_species, "Chaetodipus", "penicillatus")
plot2
```
#4
time-series plot of genus Chaetodipus and species
```{r}
library("ggplot2")
ggplot(data = plot1, mapping = aes(x = year, y = count)) + 
  geom_point() +
  labs(x = "Year", y = "Count", title = " Timed series for Genus: Dipodomys | Species: merriami")
```
#5 
time-series plot of genus Chaetodipus and species penicillatus
added color to the plots using the color func and a size of 1 using the size func
added labels using labs 
added a line that connects the plots using geom_ line func 
 
```{r}
library("ggplot2")
ggplot(data = plot2, mapping = aes(x = year, y = count)) + 
  geom_point(color = "blue", size = 1 ) +
  geom_line(color="blue" ) +
  labs(x = "Year", y = "Number of Individuals", title = " Timed series for Genus: Chaetodipus | Species: penicillatus")

```
