---
title: "functions-prep.Rmd"
author: "Azul Carrillo"
date: "2022-11-08"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Writing our own functions in r

-   allows us to understand code without understanding all of the details

```{r}
sum(c(1,2,3))
funtion_name <- function(inputs) {
  output_value <- do_something (inputs)
  return(output_value)
}

{
  a = 2 
  b = 3 
  a + b 
  
}

```

creating a function - define the instructions for a calculation haven't actually used it

```{r}
calc_shrub_vol <- function(length, width, height) {
  area <- length * width 
  volume <-  area * height 
  return(volume)
}

```

running our function

```{r}
calc_shrub_vol(0.8, 1.6, 2.0 )
shrub_vol <-  calc_shrub_vol(0.8, 1.6, 2.0) # storing the output in our function so that we can use it later 
```

2.  how to treat functions as black boxes - function should only know the inputs we pass arguments\
    the program should not Khow anything that goes in the function but the output that the function passes back down to it

```{r}
shrub_vol <-  calc_shrub_vol(0.8, 1.6, 2.0)
```
3. setting different values for arguments 
- recreating a function height=1
setting defaults to our functions 
```{r}
calc_shrub_vol <- function(length, width, height = 1) {
  area <- length * width 
  volume <-  area * height 
  return(volume)
}
calc_shrub_vol(0.8, 1.6, 2.0 )
calc_shrub_vol(0.8, 1.6) # since we gave a height a default value of 1 when the height is not specifyed itll be a value of 1

```
4. When To Use Named And Unnamed Arguments
pos based for things that are required 
named arg things ghat are optional 
only exeption is when we might get confused 
```{r}
calc_shrub_vol(length = 0.8 , width = 1.6, height = 2.0)
calc_shrub_vol(height = 2.0, length = 0.8 , width = 1.6) #order doesnt matter when named 
calc_shrub_vol(0.8 ,  1.6, height = 2.0)
```

5. combining functions 

```{r}
library(dplyr)
calc_shrub_vol <- function(length, width, height = 1) {
  area <- length * width 
  volume <-  area * height 
  return(volume)
}

est_shrub_mass <- function(volume) {
  mass <- 2.65 * volume^0.9 
  return(mass)
}
#combine the two functions 
shrub_volume <- calc_shrub_vol(0.8, 1.6, 2.0)
shrub_mass <- est_shrub_mass(shrub_vol)


#or use pipes 
shrub_mass <- calc_shrub_vol(0.8, 1.6, 2.0) %>%
  est_shrub_mass()

#nest functions together can make cose difficult to read 
shrub_mass <- est_shrub_mass(calc_shrub_vol(0.8, 1.6, 2.0))
```
6. Calling Functions Inside Of Other Functions

we should never assume that variables from the outter program are avaliable inside of a function but we can allways assume that finctions that are avaliable in the outter program are avaliable to functions to do calculations with 
```{r}
calc_shrub_vol <- function(length, width, height = 1) {
  area <- length * width 
  volume <-  area * height 
  return(volume)
}

est_shrub_mass <- function(volume) {
  mass <- 2.65 * volume^0.9 
  return(mass)
}

est_shrub_dim <- function(length, width, height = 1){
  volume <- calc_shrub_vol(length, width, height)
  mass <- est_shrub_mass(volume)
  return(mass)
}
est_shrub_dim(0.8, 1.6, height = 2.0)
```
7. tips and tricks 
navagation menu 
collapse and expang 
change settings to diferentiate between functions and variables 
```{r}
calc_shrub_vol <- function(length, width, height = 1) {
  area <- length * width 
  volume <-  area * height 
  return(volume)
}

est_shrub_mass <- function(volume) {
  mass <- 2.65 * volume^0.9 
  return(mass)
}

est_shrub_dim <- function(length, width, height = 1){
  volume <- calc_shrub_vol(length, width, height)
  mass <- est_shrub_mass(volume)
  return(mass)
}

```


