---
title: "portal-phylogenies.Rmd"
author: "Azul Carrillo"
date: "2022-10-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Newick tree format allows us to represent trees in a readable way for computers. This uses the correspondance between the trees and the nested parenthesis. In other words this format is used to describe in detail a phylogenetic tree as text. In the format the parenthesis group the sequence names. The branch lengths can be included using a colon followed by the length. Finally this is finished by a semicolon no branch length ex` (,(,,),); ` this will just add the sequence names. Ex. branch length included  ` (B:1.0,(A:2.0,C:3.0,E:4.0):5.0,D:6.0); `. 

# package 'ape'

```{r}
#install.packages("ape")
library(ape)
```

```{r}
portal_tree <- read.tree(file = "../raw-data/portal-tree.tre")
portal_tree
```
branch length Dna- will be propotional to mutation rate by some unit of time, change in DNA per kind 

```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
small_tree
```
# The structure of a tree in R 
classes we've been working with data.frames next we will work with is phylo 
```{r}
class(portal_tree)
length(portal_tree) # gives the number of 
names(portal_tree)
#portal_tree$Nnode
head(portal_tree["tip.label"])
head(portal_tree[[3]])
summary(portal_tree)
str(portal_tree) # $ to access elements
head(portal_tree$edge)
```
list is similar to a data frame but it allows combining object of any kind 

```{r}
plot.phylo(x = portal_tree)
```
An extension of gg plt for phylogenetic visualization `ggtree`
to install packages from Bioconductor repo we need a CRAN package called BiocManager
```{r}
#install.packages("BiocManager")
library(BiocManager)
#install("ggtree")
library(ggtree)
```
Plotting a Phylogenetic tree 
```{r}
ggtree(portal_tree)
```

```{r}
#this is equivalent to 
ggplot(portal_tree, aes(x,y)) +
  geom_tree() +
  theme_tree()
```
to add a skale  use the function geom_treescales():
```{r}
ggtree(portal_tree)+
  geom_treescale()


```
 
 plot the small tree of five species of primates adn include a scale what is the difference between the portal and the small tree 
```{r}
ggtree(small_tree) +
  geom_treescale()
```

```{r}
class(small_tree)
length(small_tree) # gives the number of 
names(small_tree)
small_tree$Nnode
small_tree["tip.label"]
head(small_tree[[3]])
summary(small_tree)
str(small_tree) # $ to access elements
head(small_tree$edge)
# for portal_tree the Number of tips: 43 Number of nodes: 345 therefor it is larger than the small trees  Number of tips: 5 Number of nodes: 4 edge is the branch small trees we have a branch length under edge.node 
```


```{r}
small_tree$edge.length
head(small_tree$node.label)
head(portal_tree$node.label)
```
#Add tip labels and node labels 
because a plo is nothing with no labels 
```{r}
ggtree(portal_tree) +
  geom_treescale() +
  geom_tiplab(size = 1, color= "purple", fontface = "italic")
```
add a limit to the plot so that we can see the labels 

```{r}
ggtree(portal_tree) +
  geom_treescale() +
  geom_tiplab(size = 1, color= "purple", fontface = "italic") +
  xlim(NA, 90)
```

```{r}
ggtree(small_tree) +
  geom_treescale() +
  geom_tiplab(size = 2, color= "purple", fontface = "italic") +
  xlim(NA, 2)
```

```{r}
library(ape)
branching.times(small_tree)
```
Evolution - how living things have changed (evolved)
branches aka edges or lineages 
node - event happened 
roots in a tree- represent the common ancestor to all lineage in the trees 
# creating differnt tree tepresentation 
```{r}
library(ggtree)
ggtree(portal_tree, layout = "fan", open.ange = 45) -> plot1
```

```{r}
#install.packages("aplot")
library(aplot)
plot_list(plot1, plot1, labels = c("Circular Tree of Portal Species", "The same "), tag_size = 7 )
```

```{r}
plot1.<- ggtree(portal_tree, layout="rectangular")
plot2 <-ggtree(portal_tree, layout="ape")
plot3 <-ggtree(portal_tree, layout="roundrect")
plot4 <-ggtree(portal_tree, layout="slanted")
plot5 <- ggtree(portal_tree, layout="ellipse")
plot6 <- ggtree(portal_tree, layout="dendrogram") 
plot7 <-ggtree(portal_tree, layout="circular")
plot8 <-ggtree(portal_tree, layout="radial")
plot9 <- ggtree(portal_tree, layout="fan", open.angle = 90)
plot10 <- ggtree(portal_tree, layout="equal_angle")
plot11 <-ggtree(portal_tree, layout="daylight")
plot12<- ggtree(portal_tree, layout="unrooted")
```
Part 2: 
Create a subplot for each tree.  
```{r}
plot_list(plot1., plot2,plot3, plot4,plot5, plot6, plot7, plot8,plot9,plot10, plot11,plot12, tag_levels = "A")
```
part 3: 
```{r}
my_labels<-paste(letters[1:12], "=","rooted") 

plot_list(plot1., plot2,plot3, plot4,plot5, plot6, plot7, plot8,plot9,plot10, plot11,plot12, labels = my_labels, tag_size=5)
```
j,k,l,b  un-rooted 
taxonomy - classifies
phylogeny - study of evolutionary relationships 

```{r}
taxonomy<- read.csv("../raw-data/portal-species-taxonomy.csv")
```

```{r}
write.csv(taxonomy, file ="../raw-data/portal-species-taxonomy.csv", row.names = FALSE)
nrow(taxonomy)
View(taxonomy)
```

```{r}
portal_tree
```

```{r}
library(tidytree)
library(dplyr)
full_join(portal_tree,taxonomy, by = "label")
head(taxonomy)
```

```{r}
left_join(portal_tree,taxonomy, by = "label")
#the join failed 
#we need to check our species labels  because we had all NA values 
head(taxonomy)
taxonomy$label
```
replace the tip - labels from our original portal tree with a new vector of tip 
```{r}
tip_labels <-portal_tree$tip.label
tip_labels
tip_labels_cleaned<- gsub("_"," ",tip_labels)
tip_labels_cleaned
portal_tree[["tip.label"]]<-tip_labels_cleaned
portal_tree$tip.label
```


```{r}
library(tidytree)
tree_data <- left_join(portal_tree,taxonomy, by = "label")
tree_data
class(tree_data)
```
color by a variable in our column 
```{r}
ggtree(tr= tree_data, mapping = aes(color= taxa))+
  geom_tiplab(size = 2, frontface="italic")+
  xlim(0,100)
ggsave(filename = "../fig/portal-tree-1.png")

```


```{r}
class(portal_tree)
typeof(portal_tree)
str(portal_tree)
portal_tree

```

```{r}
portal_tree <-("../raw-data/portal-tree.tre")
```

```{r}
taxonomy <- read.csv("../raw-data/portal-species-taxonomy.csv")
surveys <- read.csv("../raw-data/surveys.csv")
species <- read.csv("../raw-data/species.csv")
```

```{r}
library(ggtree)

ggtree(tr= tree_data, mapping = aes(color= taxa)) +
  geom_tiplab(size = 2.5, frontface="italic") +
  xlim(0,100)

ggsave(filename = "../fig/portal-tree-1.png")
tree_data
```
```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
branching.times(small_tree)

ggtree(small_tree) + 
  xlim(0,1.5) +
  geom_tiplab(fontface = "italic")

label <- small_tree$tip.label

taxa <- c("Hominini", "Hominidae") 
common_name <- c("Human", "Orangutan")
data.frame ("label", "taxa", "common_name")
``` 

```{r}
install.packages("TDbook")
```

```{r}
library(TDbook)
data("tree_boots")
data("df_tip_data")
data("df_inode_data")
tree_boots # 7 tips and 6 internal nodes, Rooted; includes branch lengths.
#column "Newick_label"change to "label" in  df_ tip_ data and in df_inode_data 
```

```{r}
renamed_tip_data <- df_tip_data %>%
  mutate(label = Newick_label)
colnames(renamed_tip_data )
renamed_inode_data <- df_inode_data %>%
  mutate(label = newick_label)
colnames(renamed_inode_data)
renamed_inode_data
```

```{r}
colnames(df_tip_data)[1] <- "label"
colnames(df_tip_data)
colnames(df_inode_data)[1] <- "label"
colnames(df_inode_data) 
```

```{r}
names(tree_boots)
colnames(df_tip_data)
intersect (names(tree_boots), colnames(df_tip_data))
tree_joined <- left_join(tree_boots, df_tip_data, by = "label")
tree_joined
```

```{r}
colnames(df_tip_data)
df_tip_data$rank

ggtree(tree_joined) +
  geom_tippoint(mapping = aes(shape = trophic_habit, 
                               color= trophic_habit, 
                              size = mass_in_kg))+
  geom_tiplab(fontface = "italic") + 
  xlim(0,4)

df_tip_data
```

```{r}

library(dplyr)
colnames(surveys)
surveys
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(species_id) %>%
  summarise(mean_weight = mean(weight)) -> species_weight
nrow(species_weight)
spp_weight <- inner_join(species, species_weight)
spp_weight

colnames (taxonomy) 
taxonomy_weight<- full_join(taxonomy, spp_weight)
taxonomy_weight

library(tidytree)
tree_join <- left_join(portal_tree, taxonomy_weight, by = "label")
portal_tree

tree_join
ggtree(tree_join) +
  geom_tiplab(fontface = "italic", size = 2) +
  xlim(0, 20) +
  geom_tippoint(mapping = aes(shape =  "circle", size = mean_weight, color = taxa))
```