---
title: "portal-phylogenies.Rmd"
author: "Azul Carrillo"
date: "2022-10-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Newick tree format allows us to represent trees in a readable way for computers. This uses the correspondance between the trees and the nested parenthesis. In other words this format is used to describe in detail a phylogenetic tree as text. In the format the parenthesis group the sequence names. The branch lengths can be included using a colon followed by the length. Finally this is finished by a semicolon no branch length ex` (,(,,),); ` this will just add the sequence names. Ex. branch length included  ` (B:1.0,(A:2.0,C:3.0,E:4.0):5.0,D:6.0); `. 

# package 'ape'

```{r}
#install.packages("ape")
library(ape)
```

```{r}
portal_tree <-read.tree("../raw-data/portal-tree.tre")
```

```{r}
taxonomy <- read.csv("../raw-data/portal-species-taxonomy.csv")
surveys <- read.csv("../raw-data/surveys.csv")
species <- read.csv("../raw-data/species.csv")
```

branch length Dna- will be propotional to mutation rate by some unit of time, change in DNA per kind 

```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
small_tree
```
# The structure of a tree in R 
classes we've been working with data.frames next we will work with is phylo 
```{r}
class(portal_tree)
length(portal_tree) # gives the number of 
names(portal_tree)
#portal_tree$Nnode
head(portal_tree["tip.label"])
head(portal_tree)
summary(portal_tree)
str(portal_tree) # $ to access elements
head(portal_tree)
```
list is similar to a data frame but it allows combining object of any kind 

```{r}
#plot.phylo(x = portal_tree)
```
An extension of gg plt for phylogenetic visualization `ggtree`
to install packages from Bioconductor repo we need a CRAN package called BiocManager
```{r}
#install.packages("BiocManager")
library(BiocManager)
#install("ggtree")
library(ggtree)
```
Plotting a Phylogenetic tree 
```{r}
library(ggtree)
portal_tree <-read.tree("../raw-data/portal-tree.tre")
ggtree(portal_tree)
```

```{r}
#this is equivalent to 
ggplot(portal_tree, aes(x,y)) +
  geom_tree() +
  theme_tree()
```
to add a skale  use the function geom_treescales():
```{r}
ggtree(portal_tree)+
  geom_treescale()


```
 
 plot the small tree of five species of primates adn include a scale what is the difference between the portal and the small tree 
```{r}
ggtree(small_tree) +
  geom_treescale()
```

```{r}
class(small_tree)
length(small_tree) # gives the number of 
names(small_tree)
small_tree$Nnode
small_tree["tip.label"]
head(small_tree[[3]])
summary(small_tree)
str(small_tree) # $ to access elements
head(small_tree$edge)
# for portal_tree the Number of tips: 43 Number of nodes: 345 therefor it is larger than the small trees  Number of tips: 5 Number of nodes: 4 edge is the branch small trees we have a branch length under edge.node 
```


```{r}
small_tree$edge.length
head(small_tree$node.label)
head(portal_tree$node.label)
```
#Add tip labels and node labels 
because a plo is nothing with no labels 
```{r}
ggtree(portal_tree) +
  geom_treescale() +
  geom_tiplab(size = 1, color= "purple", fontface = "italic")
```
add a limit to the plot so that we can see the labels 

```{r}
ggtree(portal_tree) +
  geom_treescale() +
  geom_tiplab(size = 1, color= "purple", fontface = "italic") +
  xlim(NA, 20)
```

```{r}
ggtree(small_tree) +
  geom_treescale() +
  geom_tiplab(size = 2, color= "purple", fontface = "italic") +
  xlim(NA, 2)
```

```{r}
library(ape)
branching.times(small_tree)
```
Evolution - how living things have changed (evolved)
branches aka edges or lineages 
node - event happened 
roots in a tree- represent the common ancestor to all lineage in the trees 
# creating differnt tree tepresentation 
```{r}
library(ggtree)
ggtree(portal_tree, layout = "fan", open.ange = 45) -> plot1
```

```{r}
#install.packages("aplot")
library(aplot)
plot_list(plot1, plot1, labels = c("Circular Tree of Portal Species", "The same "), tag_size = 7 )
```

```{r}
plot1.<- ggtree(portal_tree, layout="rectangular")
plot2 <-ggtree(portal_tree, layout="ape")
plot3 <-ggtree(portal_tree, layout="roundrect")
plot4 <-ggtree(portal_tree, layout="slanted")
plot5 <- ggtree(portal_tree, layout="ellipse")
plot6 <- ggtree(portal_tree, layout="dendrogram") 
plot7 <-ggtree(portal_tree, layout="circular")
plot8 <-ggtree(portal_tree, layout="radial")
plot9 <- ggtree(portal_tree, layout="fan", open.angle = 90)
plot10 <- ggtree(portal_tree, layout="equal_angle")
plot11 <-ggtree(portal_tree, layout="daylight")
plot12<- ggtree(portal_tree, layout="unrooted")
```
Part 2: 
Create a subplot for each tree.  
```{r}
plot_list(plot1., plot2,plot3, plot4,plot5, plot6, plot7, plot8,plot9,plot10, plot11,plot12, tag_levels = "A")
```
part 3: 
```{r}
my_labels<-paste(letters[1:12], "=","rooted") 

plot_list(plot1., plot2,plot3, plot4,plot5, plot6, plot7, plot8,plot9,plot10, plot11,plot12, labels = my_labels, tag_size=5)
```
j,k,l,b  un-rooted 
taxonomy - classifies
phylogeny - study of evolutionary relationships 

```{r}
taxonomy<- read.csv("../raw-data/portal-species-taxonomy.csv")
```

```{r}
write.csv(taxonomy, file ="../raw-data/portal-species-taxonomy.csv", row.names = FALSE)
nrow(taxonomy)

```

```{r}
portal_tree
```

```{r}
library(tidytree)
library(dplyr)
full_join(portal_tree,taxonomy, by = "label")
head(taxonomy)
```

```{r}
left_join(portal_tree,taxonomy, by = "label")
#the join failed 
#we need to check our species labels  because we had all NA values 
head(taxonomy)
taxonomy$label
```
replace the tip - labels from our original portal tree with a new vector of tip 
```{r}
tip_labels <-portal_tree$tip.label
tip_labels
tip_labels_cleaned<- gsub("_"," ",tip_labels)
tip_labels_cleaned
portal_tree[["tip.label"]]<-tip_labels_cleaned
portal_tree$tip.label
```
```{r}
portal_tree <- read.tree("../raw-data/portal-tree.tre")

```

```{r}
taxonomy <- read.csv("../raw-data/portal-species-taxonomy.csv")
```

```{r}
library(tidytree)
tree_data <- left_join(portal_tree,taxonomy, by = "label")
tree_data
class(tree_data)
```
color by a variable in our column 
```{r}
ggtree(tr= tree_data, mapping = aes(color= taxa))+
  geom_tiplab(size = 2, frontface="italic")+
  xlim(0,100)
ggsave(filename = "../visuals/portal-tree-1.png")

```


```{r}
class(portal_tree)
typeof(portal_tree)
str(portal_tree)
portal_tree

```

```{r}
portal_tree <-read.tree("../raw-data/portal-tree.tre")
```

```{r}
taxonomy <- read.csv("../raw-data/portal-species-taxonomy.csv")
surveys <- read.csv("../raw-data/surveys.csv")
species <- read.csv("../raw-data/species.csv")
```

```{r}
library(ggtree)

ggtree(tr= tree_data, mapping = aes(color= taxa)) +
  geom_tiplab(size = 2.5, frontface="italic") +
  xlim(0,100)

ggsave(filename = "../visuals/portal-tree-1.png")
tree_data
```
Ex for the hw 
```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
branching.times(small_tree)

ggtree(small_tree) + 
  xlim(0,1.5) +
  geom_tiplab(fontface = "italic")

label <- small_tree$tip.label[1:2]

taxa <- c("Hominini", "Hominidae") 
common_name <- c("Human", "Orangutan")

data.frame(label, taxa, common_name)
``` 

```{r}
#install.packages("TDbook")
```

```{r}
library(TDbook)
data("tree_boots")
data("df_tip_data")
data("df_inode_data")
tree_boots # 7 tips and 6 internal nodes, Rooted; includes branch lengths.
#column "Newick_label"change to "label" in  df_ tip_ data and in df_inode_data 
```

```{r}
renamed_tip_data <- df_tip_data %>%
  mutate(label = Newick_label)
colnames(renamed_tip_data )
renamed_inode_data <- df_inode_data %>%
  mutate(label = newick_label)
colnames(renamed_inode_data)
renamed_inode_data
```

```{r}
colnames(df_tip_data)[1] <- "label"
colnames(df_tip_data)
colnames(df_inode_data)[1] <- "label"
colnames(df_inode_data) 
```

```{r}
names(tree_boots)
colnames(df_tip_data)
intersect (names(tree_boots), colnames(df_tip_data))
tree_joined <- left_join(tree_boots, df_tip_data, by = "label")
tree_joined
```

```{r}
colnames(df_tip_data)
df_tip_data$rank

ggtree(tree_joined) +
  geom_tippoint(mapping = aes(shape = trophic_habit, 
                               color= trophic_habit, 
                              size = mass_in_kg))+
  geom_tiplab(fontface = "italic") + 
  xlim(0,4)

df_tip_data
```

```{r}

library(dplyr)
colnames(surveys)
surveys
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(species_id) %>%
  summarise(mean_weight = mean(weight)) -> species_weight
nrow(species_weight)
spp_weight <- inner_join(species, species_weight)
spp_weight

colnames (taxonomy) 
taxonomy_weight<- full_join(taxonomy, spp_weight)
taxonomy_weight

library(tidytree)
tree_join <- left_join(portal_tree, taxonomy_weight, by = "label")
portal_tree

tree_join
ggtree(tree_join) +
  geom_tiplab(fontface = "italic", size = 2) +
  xlim(0, 20) +
  geom_tippoint(mapping = aes(shape =  "circle", size = mean_weight, color = taxa))
getwd()
```
Homework for 11/03/2022
```{r}
#1 here i obtained the average weight and hindfoot length and filtered all na values 
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(species_id) %>%
  summarise(mean_weight = mean(weight)) -> species_weight

surveys %>%
  filter(!is.na(hindfoot_length))%>%
  group_by(species_id) %>%
  summarise(mean_hindfoot_length =mean(hindfoot_length)) -> species_hindfoot_length

#2  Create a new data frame that contains the taxonomy data plus the averaged data per species that you got on last question.
average_data <- full_join(species_weight, species_hindfoot_length)
head(average_data)
#joined both averages and the taxonomy 
taxonomy_average<- full_join(taxonomy, average_data)
taxonomy_average

# joined the tree and the data frame with the taxonomy and the averages 
joined_data  <- left_join(portal_tree, taxonomy_average, by = "label")

joined_data

#3 Create two plots with data on the tips, one with the average weight and the other with average hindfoot length. Make sure to also add tip labels, formatted in italics.
ggtree(joined_data) +
  geom_tiplab(fontface = "italic", size = 2) +
  xlim(0, 20) +
  geom_tippoint(mapping = aes(shape =  "circle", size = mean_weight, color = taxa))
```

```{r}
#plot 2 with the mean hindfoot 
ggtree(joined_data) +
  geom_tiplab(fontface = "italic", size = 2) +
  xlim(0, 20) +
  geom_tippoint(mapping = aes(shape =  "circle", size = mean_hindfoot_length, color = taxa))
```
Homework for 10/27/2022
#A taxonomy table for small_tree
1. Find the appropriate scientific group labels for each genus in small_tree using this tree as guide
2. Create a data frame with 3 columns:
- a "label" column with the names of the tip labels of small_tree.
- a "taxa" column with the scientific names of the group that each genus belongs to.
- a "common_name" column with the common names of the group that each genus belongs to.
3. Join your tree and your table
4. Create two different tree plots using taxa and common name to color the tips of the tree.

```{r}
#1
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
#2 creating a data frame with 3 columns 
label <- small_tree$tip.label
taxa <- c("Hominini", "Hominidae","Macaques","Atelidae", "Galagidae")
common_name <- c("Human", "Orangutan","Macaque","Spider Monkey", "Galagos")
#data frame is created
Etymology <- data.frame(label, taxa, common_name)

#3joining the tree and the data frame by label using left_join
full_join  <- left_join(small_tree, Etymology, by = "label")

```
#4 Create two different tree plots using taxa and common name to color the tips of the tree.
```{r}
#4.1
ggtree(full_join) +
  geom_tiplab(fontface = "italic", size = 5) +
  xlim(0, 2) +
  geom_tippoint(mapping = aes(shape =  "circle",size = taxa, color = taxa))
```

```{r}
#4.2
ggtree(full_join) +
  geom_tiplab(fontface = "italic", size = 5) +
  xlim(0, 2) +
  geom_tippoint(mapping = aes(shape =  "circle",size = common_name, color = common_name))
```