---
title: "Node labels on trees"
author: "Azul Carrillo"
date: "2022-11-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Data set the we are working with comes with thw TDbook r package

load the data packages 

```{r}
library("TDbook")
library(ggtree)
library(tidytree)


```
load the data from the data package suing the `data()`
the data we are loading 
-tree_boots 
- df_tip_data 
-df_ inside_data 

Exercise
1. What is the class of this three objects
2. How many elements does `tree_boots` has
3. How many rows does `df_tip_data` and `df_inode_data` have.
4. Compare the row numbers of both objects to the lengths of the `$tip.label` and `$node.label` elements of `tree_boots`.
```{r}
data("tree_boots", "df_inode_data", "df_tip_data")

tree_boots
class(tree_boots) #phylo  7 tips and 6 internal nodes.Rooted; includes branch lengths.

df_inode_data
class(df_inode_data) #data frame 6 rows 6 col 
nrow(df_inode_data)
length(df_inode_data)


df_tip_data
class(df_tip_data) #data frame 7 rows 10 col 
nrow(df_tip_data)
length(df_tip_data)
#joining data frameas and the tree 
library(tidytree)
colnames(df_bar_data)
df_tip_data$Newick_label
tree_boots$tip.label
#rename Newick_ label
df_mutate<- mutate(df_tip_data, label = Newick_label)
colnames(df_mutate)
tree_joined <- left_join(tree_boots, df_mutate, by = "label")


df_tip_data

colnames(df_tip_data)-> my_colnames
my_colnames
typeof(my_colnames)#"character"
is.vector(my_colnames)
my_colnames[1]<- "column 1 "
my_colnames[5]
my_colnames
my_colnames[10] <- "taxonomy"

colnames(df_tip_data)
colnames(df_tip_data)[1]<-"label" 
colnames(df_tip_data)
tree_joined <- left_join(tree_boots, df_tip_data)

tree_joined
tree_boots

``` 

```{r}
tree_joined@phylo
tree_joined@extraInfo
```
##plot node labels 
```{r}
df_tip_data
df_inode_data

```

```{r}
colnames(df_inode_data) [1]<-"label" 
tree_data2 <- left_join(tree_boots, df_inode_data)
tree_data2
```

```{r}
ggtree(tree_data2)+
  geom_tiplab(fontface = "italic")+
  xlim(0,5) +
  geom_nodelab(mapping = aes(label = vernacularName, color = vernacularName))
  
  
```


```{r}
portal_tree <-("../raw-data/portal-tree.tre")
```

```{r}
taxonomy <- read.csv("../raw-data/portal-species-taxonomy.csv")
surveys <- read.csv("../raw-data/surveys.csv")
species <- read.csv("../raw-data/species.csv")
```

```{r}
library(magrittr)
library(dplyr)
library(ggtree)
library(tidytree)
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(species_id) %>%
  summarise(mean_weight = mean(weight)) -> species_weight

surveys %>%
  filter(!is.na(hindfoot_length))%>%
  group_by(species_id) %>%
  summarise(mean_hindfoot_length =mean(hindfoot_length)) -> species_hindfoot_length

average_data <- full_join(species_weight, species_hindfoot_length)
head(average_data)
#joined both averages and the taxonomy 
taxonomy_average<- full_join(taxonomy, average_data)
colnames(taxonomy_average)
library(dplyr)
portal_tree <-read.tree("../raw-data/portal-tree.tre")
portal_joined <- left_join(portal_tree, taxonomy_average)
```


```{r}
colnames(portal_joined)

ggtree(portal_joined)+
  geom_tiplab( fontface = "italic",size= 4) +
  xlim(0,20) +
  geom_nodelab()
```