---
title: "review"
author: "Azul Carrillo"
date: "2022-11-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```
#1 Write a conditional statement that checks if surveys.csv exists in the working directory, if it doesn’t then downloads it from https://ndownloader.figshare.com/files/2292172 using download.file(), and finally loads the file into a data frame and displays the first few rows using the head() function. The url needs to be in quotes since it is character data.
```{r}
getwd()
"surveys.csv" == c(list.files("/Users/atziri/Bio 195-197/Data Science/raw-data"))

surveys <- list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") == "surveys.csv"


is.element("surveys.csv", list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") )
```

```{r}
```

```{r}
is.element("surveys.csv", list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") )

```

```{r}

surveys <- list.files("../raw-data")

if (is.element("surveys.csv", surveys)){
  print("file is downloaded")
} else {
   print("file is not downloaded") 
  download.file("https://ndownloader.figshare.com/files/2292172",
              "../raw-data/surveys-download.csv")
surveys_data <- read.csv("../raw-data/surveys-download.csv")
 head(surveys_data)
}


```
#2 Make a version of this conditional statement that is a function, where the name of the file is the first argument and the link for downloading the file is the second argumen
```{r}
#This function tests if a file is in the raw-data directory and if not it download it and read it as data frame with the read, csv function:

reading_csv <- function(file_name, file_link) {

# 1. test if file_name is in the raw-data folder
# file_name <- "species.csv"
  
test <- !is.element(file_name, list.files(path = "../raw-data"))

# 2. if test is FALSE, download the file

if (test) {
  # Option 1: save it with a random name:
  # download.file(url = file_link, destfile = "../raw-data/temporary.csv")
  
  # result <- read.csv(file = "../raw-data/temporary.csv")

# Option 2: save it with the name given in file nome:
  destination_file <- stringr::str_c("../raw-data/", file_name) 
  download.file(url = file_link, destfile = destination_file)
  result <- read.csv(file = destination_file)
}
  return(result)
}
```

```{r}

reading_csv <- function(file_name, file_link) {
  test <- is.element(file_name, list.files(path = "../raw-data"))# removed the exclamation mark for it to show the table when i already have a species file downloaded 
  if (test) {
    destination_file <- stringr:: str_c("../raw-data/", file_name) 
    download.file(url = file_link, destfile = destination_file)
    
    result <- read.csv(file = destination_file)
  }
   return(result)
}
```

```{r}
reading_csv(file_name = "species.csv", 
            file_link = "https://ndownloade.figshare.com/files/3299483")
```






















