---
title: "review"
author: "Azul Carrillo"
date: "2022-11-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```
#1 Write a conditional statement that checks if surveys.csv exists in the working directory, if it doesn’t then downloads it from https://ndownloader.figshare.com/files/2292172 using download.file(), and finally loads the file into a data frame and displays the first few rows using the head() function. The url needs to be in quotes since it is character data.
```{r}
getwd()
"surveys.csv" == c(list.files("/Users/atziri/Bio 195-197/Data Science/raw-data"))

surveys <- list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") == "surveys.csv"


is.element("surveys.csv", list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") )
```

```{r}
```

```{r}
is.element("surveys.csv", list.files("/Users/atziri/Bio 195-197/Data Science/raw-data") )

```

```{r}

surveys <- list.files("../raw-data")

if (is.element("surveys.csv", surveys)){
  print("file is downloaded")
} else {
   print("file is not downloaded") 
  download.file("https://ndownloader.figshare.com/files/2292172",
              "../raw-data/surveys-download.csv")
surveys_data <- read.csv("../raw-data/surveys-download.csv")
 head(surveys_data)
}


```
#2 Make a version of this conditional statement that is a function, where the name of the file is the first argument and the link for downloading the file is the second argumen
```{r}
#This function tests if a file is in the raw-data directory and if not it download it and read it as data frame with the read, csv function:

reading_csv <- function(file_name, file_link) {

# 1. test if file_name is in the raw-data folder
# file_name <- "species.csv"
  
test <- !is.element(file_name, list.files(path = "../raw-data"))

# 2. if test is FALSE, download the file

if (test) {
  # Option 1: save it with a random name:
  # download.file(url = file_link, destfile = "../raw-data/temporary.csv")
  
  # result <- read.csv(file = "../raw-data/temporary.csv")

# Option 2: save it with the name given in file nome:
  destination_file <- stringr::str_c("../raw-data/", file_name) 
  download.file(url = file_link, destfile = destination_file)
  result <- read.csv(file = destination_file)
}
  return(result)
}
```

```{r}

reading_csv <- function(file_name, file_link) {
  test <- is.element(file_name, list.files(path = "../raw-data"))# removed the exclamation mark for it to show the table when i already have a species file downloaded 
  if (test) {
    destination_file <- stringr:: str_c("../raw-data/", file_name) 
    download.file(url = file_link, destfile = destination_file)
    
    result <- read.csv(file = destination_file)
  }
   return(result)
}
```

```{r}
reading_csv(file_name = "species.csv", 
            file_link = "https://ndownloade.figshare.com/files/3299483")
```
#Exercise 2: Multi-file Analysis

##1. If individual_collar_data.zip is not already in your working directory download the zip file using download.file()

##2. Unzip it using unzip()

##3. Obtain a list of all of the files with file names matching the pattern "collar-data-.*.txt" (using list.files())

```{r}
library(dplyr)
library(ggplot2)
download.file("http://www.datacarpentry.org/semester-biology/data/individual_collar_data.zip", "collar.zip")
unzip("collar.zip", exdir = "../raw-data")
collar_data_files <- list.files(pattern = "collar-")
collar_data_files

```

##4. Use a loop to load each of these files into R and make a line plot (using geom_path()) for each file with long on the x axis and lat on the y axis. Graphs, like other types of output, won’t display inside a loop unless you explicitly display them, so you need put your ggplot() command inside a print() statement. Include the name of the file in the graph as the graph title using labs().

##5. Add code to the loop to calculate the minimum and maximum latitude in the file, and store these values, along with the name of the file, in a data frame. Show the data frame as output.
```{r}
min_results <- vector(mode = "integer", length(collar_data_files)) 
max_results <- vector(mode = "integer", length(collar_data_files)) 


for(i in 1:length(collar_data_files)){
  filename <- collar_data_files[i]
  data <- read.csv(filename) 
  data %>%
    ggplot(aes(x = long, y = lat )) +
    labs(title = filename) +
  geom_path() -> plots
    print(plots) 
   min_results[i] <- min(data$lat)
   max_results[i] <- max(data$lat)
}

```
```{r}

min_max_data <- data.frame(collar_data_files, min_results, max_results)

min_max_data
```

##Solution 2 
```{r}
all_min <-  vector()
all_max <- vector()
length(all_min)
file_name <- vector()

for(i in collar_data_files){
  print(i)
  print(getwd())
  file_name <- c(file_name,i)
  collar_data_table <- read.csv(file = i )
  min_lat <- min (collar_data_table[,"lat"])
  all_min <- c(all_min,min_lat)
  max_lat <- max(collar_data_table[,"lat"])
  all_max <- c(all_max, max_lat)
  print(all_min)
  print(all_max)
}

min_max_dataframe <- data_frame(file_name, all_min, all_max)
min_max_dataframe



```
```{r}
#solution 3 
library(stringr)
all_min_lat <- vector(mode = "integer", length = length(collar_data_files)) 
all_max_lat <- all_min_lat
all_file_names <- all_min_lat
length(all_min)
length(all_max)
all_file_names

for (i in 1:length(collar_data_files)){
  file_name_and_path <- str_c("../raw-data/",collar_data_files[i])
  all_file_names[i] <- file_name_and_path
  print(file_name_and_path)
  collar_data_table <- read.csv(file = file_name_and_path)
  min_lat <- min(collar_data_table$lat)
  all_min_lat[i] <- min_lat
  max_lat <- max(collar_data_table$lat)
  all_max_lat[i] <- max_lat
  print(all_min_lat)
  print(all_max_lat)
}

min_max_dataframee <- data_frame(all_file_names, all_min_lat, all_max_lat)
min_max_dataframee

```
#Exercise 3: A function for the UHURU data set


##1. Explain what each line of code in the body of the function is doing. Add the explanations to your Rmd file as comments, before each line of code.

```{r}

report_rsquared <- function(data, species, formula){
  subset <- dplyr::filter(data, ANT == species)#subset the ANT row name by a logical condition changing it to species 
  test <- lm(formula, data = subset) #lm fun stands for linear models is used to fit linear models to data frames 
  rsquared <- round(summary(test)$r.squared, 3)#round function rounds off values to a specific number of decimal value
  output <- data.frame(species = species, r2 = rsquared)#creates a data frame from the arguments given and in order for this to be printed at the end of the func it was created into an object 
  return(output)
}
```

##2. Execute the function using the UHURU data and specifying species = "CM" and formula = "AXIS1~CIRC".
```{r}
UHURU_data <- read.csv(file = "../raw-data/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
UHURU_data
report_rsquared(UHURU_data,"CM","AXIS1~CIRC")
```

##3. Modify the function so that it also determines if() the rsquared is significant based on a given threshold. The modified function should return() the species, rsquared and a significance value of "S" for a relationship with an rsquared > threshold or "NS" for an rsquared < threshold.
```{r}
report_rsquared_modified <- function(data, species, formula, threshold){
  subset <- dplyr::filter(data, ANT == species)
  test <- lm(formula, data = subset)
  rsquared <- round(summary(test)$r.squared, 3)
   if (rsquared > threshold ){
     print("S")
} else if (rsquared < threshold)
   print("NS") 
  output <- data.frame(species = species, r2 = rsquared)
  return(output)
}

```

##4. Execute your modified function for species of "CM", "CS", and "TP" given a threshold = 0.667.
```{r}
report_rsquared_modified(UHURU_data,"CM","AXIS1~CIRC", 0.667)
report_rsquared_modified(UHURU_data,"CS","AXIS1~CIRC", 0.667)
report_rsquared_modified(UHURU_data,"TP","AXIS1~CIRC", 0.667)
```

#Exercise 4: Multi-file Analysis DNA or RNA Iteration


Write a function, dna_or_rna(sequence), that determines if a sequence of base pairs is DNA, RNA, or if it is not possible to tell given the sequence provided. Since all the function will know about the material is the sequence the only way to tell the difference between DNA and RNA is that RNA has the base Uracil ("u") instead of the base Thymine ("t"). Have the function return one of three outputs: "DNA", "RNA", or "UNKNOWN".

##1. Use the function and a for loop to print the type of the sequences in the following list.
##2.Use the function and sapply to print the type of the sequences in the following list.

```{r}
library(dplyr)
library(stringr)
sequences <- c("ttgaatgccttacaactgatcattacacaggcggcatgaagcaaaaatatactgtgaaccaatgcaggcg", "gauuauuccccacaaagggagugggauuaggagcugcaucauuuacaagagcagaauguuucaaaugcau", "gaaagcaagaaaaggcaggcgaggaagggaagaagggggggaaacc", "guuuccuacaguauuugaugagaaugagaguuuacuccuggaagauaauauuagaauguuuacaacugcaccugaucagguggauaaggaagaugaagacu", "gauaaggaagaugaagacuuucaggaaucuaauaaaaugcacuccaugaauggauucauguaugggaaucagccggguc")

dna_or_rna <- function(sequences){
  if (str_detect(sequences, "u")) {
 result <-  print("RNA")
   } else if (str_detect(sequences, "t")) {
   result <- print("DNA")
   } else { 
 result <- print("UNKNOWN")
  }
     return(result)
}
sapply(sequences, dna_or_rna)

```
#Exercise 5:Energy Conversion Challenge


1. Write a function with the form convert_energy_units(energy_value, input_unit, output_unit) to convert units between the following energy values:

Joules(J),
Kilojoules(KJ),
Calories(CAL),
and Kilocalories (KCAL; this unit is used for labeling the amount of energy contained in food).
To write the equations to convert between units, consider the following:

A Kilojoule is 1000 Joules,
a Calorie is 4.1868 Joules,
a Kilocalorie is 4186.8 Joules.

```{r}
convert_energy_units <- function(energy_value, input_unit, output_unit){
   if (str_detect(input_unit, "KJ")) {
     J <- energy_value * 1000
   } else if (str_detect(input_unit, "CAL")) {
     J <- energy_value * 4.1868 
   } else {
      (str_detect(input_unit, "KCAL"))
     J <- energy_value * 4186.8
  }
     return(J)
}
```
An example of a call to this function would look like:

  energy_in_cal <- 200
  energy_in_j <- convert_energy_units(energy_in_cal, "CAL", "J")
  
2. Test your function by running the example call above.
```{r}
energy_in_cal <- 200
energy_in_j <- convert_energy_units(energy_in_cal, "CAL", "J")
energy_in_j
```
3. If either the input unit or the output unit do not match the four types given above, have the function print - “Sorry, I don’t know how to convert “ + the name of the unit provided.
```{r}
convert_energy_units <- function(energy_value, input_unit, output_unit){
   if (str_detect(input_unit, "KJ")) {
     J <- energy_value * 1000
   } else if (input_unit == "CAL") {
     J <- energy_value * 4.1868 
     print("CAL")
   } else if (input_unit == "KCAL") {
     J <- energy_value * 4186.8
     print("this should be KCAL")
   } else { 
   message("Sorry, I don’t know how to convert", input_unit)
     return(NA)
   }
     return(J)
}


```
```{r}
convert_to_J <- function(energy_value, input_unit){
   if (str_detect(input_unit, "KJ")) {
     J <- energy_value * 1000
   } else if (input_unit == "CAL") {
     J <- energy_value * 4.1868 
     message("input units are CAL")
   } else if (input_unit == "KCAL") {
     J <- energy_value * 4186.8
     message("input units are KCAL")
   } else { 
   message("Sorry, I don’t know how to convert", input_unit)
    # return(NA)
     J <- NA
   }
     return(J)
}

```
4. Use your function to answer the following questions:

a) What is the daily metabolic energy used by a human (~2500 KCALs) in Joules.
```{r}
convert_energy_units(2500, "CAL", "J")
```
```{r}
convert_energy_units(2500, "KCAL", "J")
```
b) How many times more energy does a common seal use than a human? The common seal uses ~52,500 KJ/day (Nagy et al. 1999). Use the daily human metabolic cost calculated in (4a).
```{r}
convert_energy_units(52500, "KCAL", "J")
```
c) How many ergs (ERG) are there in one kilocalorie. Since we didn’t include the erg conversion this should trigger our ‘don’t know how to convert’ message.
```{r}
convert_energy_units(1, "ERG", "J")
```
5. Make the function more efficient, and instead of writing an individual conversion between each of the different units (which requires 12 if statements) you could choose to convert all of the input units to a common scale and then convert from that common scale to the output units. This approach is especially useful if you need to add new units later.
```{r}
convert_energy_efficient <- function(energy_value, input_unit, output_unit){
  #convert input to J 
 energy_to_J <- convert_energy_units(energy_value, input_unit)
  #convert J to output unit 
  message("output unit is", output_unit)
    if (str_detect(output_unit, "KJ")) {
     energy <- energy_to_J / 1000
   } else if (output_unit == "CAL") {
     energy <- energy_to_J / 4.1868 
     print("CAL")
   } else if (output_unit == "KCAL") {
     energy <- energy_to_J / 4186.8
     print(str_c("this should be KCAL"))
   } else { 
     message("Sorry, I don’t know how to convert", input_unit)
     print("sorry again")
     return(NA)
   }

  return(energy)
}
```


```{r}
convert_energy_efficient(200, "Azul","End of the semester!" )
```

```{r}
convert_energy_efficient(1200, "CAL",output_unit = "KCAL" )
```


#thank you for your patience luna i loved having you a an instructor truly have seen progress in myself can't wait for you to become a professor. Share the news when you do!!!!  <3





