---
title: "small-mammals"
output:
  pdf_document: default
  html_document: default
date: '2022-10-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
For all the data manipulation we are going to use the 'dplyr' package 

to install a package we use the function install_packages()
but to use the function inside a package we need to load them to our environment we do this with the function 'library()'
```{r}
library(dplyr)
```


#Read teh data table 

```{r cars}
#getwd()
surveys <- read.csv(file = "../raw-data/surveys.csv")
head(surveys)
#View(surveys)
plots <- read.csv(file = "../raw-data/plots.csv")
head(plots)
species <- read.csv(file = "../raw-data/species.csv")
head(species)
```

# Subst a data table by columns 

Using the dplyr function 'subset()':

```{r}
select(surveys, year, species_id, weight)
surveys_short <- select(surveys, year, species_id, weight )
head(surveys_short)
```
Other ways to subset a data table by columns: 

```{r}
columns2keep <- c("year", "species_id", "weight")
surveys[,columns2keep]
```
## Add columns to our data 

we can use the function `mutate()`

Add an extra columb that has weight divided by 10 

```{r}
colnames(surveys)
surveys$weight_decimal <- surveys$weight/10

colnames(surveys)
```

we can use function `mutate()`
```{r}
surveys <- mutate(surveys,weight_decimal_2 =weight/10)
```

```{r}
surveys
```

```{r}
arrange(surveys, weight)
```

to arrange in a descending order : 
```{r}
arrange(surveys, desc(weight))
```


##Filter Data 

to filter unwanted data or by some specific value in our data we use the function `filter()`
! meand 
```{r}
filter(surveys, species_id == "DS")
filter(surveys, species_id !="DS")
```

We can also filter by multiple conditions : 

```{r}
temp_table <- filter(surveys, species_id <="DS", year < 1998)
tail(temp_table, 100)
```

the logical operator `&` `|` mean "AND" and "OR"
```{r}
filter(surveys, species_id == "DS" & year >= 1998)
```

```{r}
filter(surveys, species_id == "DS" | year >= 1998)
```

lets filter for values that are NA we need the function `is.na()`
```{r}
is.na(surveys$weight)
surveys$weight
filter(surveys, is.na(weight))

```

```{r}
is.na(surveys$weight)
surveys$weight
filter(surveys, !is.na(weight))
```

#Exercise 1: Data manipulation

```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
```
```{r}
subset1 <- select(surveys, year, month, day, species_id)
```
```{r}

surveys_subset <- filter(surveys, !is.na(weight))
nrow(surveys_subset)

surveys_subset <- mutate(surveys_subset, weight_kg = weight/1000)
nrow(surveys_subset)

head(surveys_subset)
surveys_subset <-select(surveys_subset, year, species_id, weight_kg)
surveys_subset_filtered <- filter(surveys_subset, species_id == "SH")
nrow(surveys_subset_filtered)
```
## Pipes 

`%>%`- `magrittr` package required 
`|>`inbuilt in R very new so we need the updated version of r 
```{r}
library(dplyr)
mean(x = c(1, 3, 65, 38, 100)) ##sequential code 
c(1, 3, 65, 38, 100) %>% mean()


sqrt(mean(x = c(1, 3, 65, 38, 100 )))
c(1, 3, 65, 38, 100) %>% mean() %>% sqrt()
```
```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
```
```{r}
surveys1 <- surveys %>%
  mutate(weight_kg = weight/1000 ) %>%
   select(year,species_id, weight_kg ) %>%
  filter(!is.na(weight_kg))
colnames(surveys1)
is.na(surveys1$weight_kg) %>% any() 

 
```


```{r}


surveys2 <- surveys %>%
   select(year, month, day, species_id, weight) %>%
  filter(species_id == "SH") %>%
  filter(!is.na(weight)) %>% 
  select(year, month, day, species_id) 

   

surveys2$species_id == "SH"

colnames(surveys2)
  colnames(surveys2)
nrow(surveys2)
```

```{r}
ds_data <-filter(surveys, species_id == "DS",!is.na(weight)) %>%
 arrange(., year) %>%
 select(., year, weight)
```

```{r}
ds_data <- filter(surveys, species_id == "DS") %>% 
  filter(!is.na(weight)) 
ds_data_by_year <- arrange(ds_data, year)
ds_weight_by_year <- select(ds_data, year, weight, species_id)
```

```{r}  
ds_weight_by_year <-surveys %>% 
  filter(species_id == "DS") %>% 
  filter(!is.na(weight)) %>% 
  arrange(year) %>% 
  select(year, weight, species_id)
nrow(ds_weight_by_year)
```

## Data Aggregation 
```{r}
surveys_by_year <- group_by(surveys, year, plot_id)
surveys_by_year
```
```{r}
plot_year_counts <- summarize(surveys_by_year, abundance = n())
```

```{r}  
plot_year_counts<-surveys %>% 
  group_by(plot_id, year) %>% 
  summarise(abundance = n())
```

Homework 
1
```{r}
surveys_individuals <- surveys %>%
  group_by(species_id) %>%
  summarize(abundance = n())

```
2
```{r}
surveys_individuals_year <- surveys %>%
  group_by(species_id, year ) %>%
  summarize(abundance = n())
```

3
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  filter(species_id == "DO") %>%
  group_by(year) %>%
  summarise(mean_year = mean(weight))
  

```



```{r}
inner_join(surveys, species, by = "species_id")
left_join(surveys, species,  by = "species_id")
right_join(surveys, species,  by = "species_id")
full_join(surveys, species,  by = "species_id")
```


how to find out if we have a column that is matching in two tables 
```{r}
surveys_colnames <- colnames(surveys)
species_colnames <- colnames(species)
surveys_colnames[7]
to_keep <-  c(1, 4, 3, 2)
to_keep <- c(FALSE, TRUE, TRUE, FALSE)
species_colnames[to_keep]

```

The %in% operator 
```{r}
1 %in% c(1,2,3)
c(1,2,3) %in% 1

matching_colnames <- colnames(species) %in% colnames(surveys)
matching_colnames
colnames(species)[matching_colnames]
```
 
 How to join more than 2 tables 
```{r}
combined <- inner_join(surveys, species, by = "species_id")
combined_full <- inner_join(combined, plots, by = "plot_id")
head(combined)
head(combined_full)
```

how do we run the cone above using a pipe

```{r}
combined_full <- surveys %>%
  inner_join( species, by = "species_id") %>%
  inner_join( plots, by = "plot_id")
head(combined_full)
  
  
```

```{r}
 surveys %>%
  inner_join(plots, by = "plot_id") %>%
   filter( plot_type == "Control" )

 
 filter(plots, plot_type == "Control" )%>%
 inner_join(surveys, by = "plot_id" ) 
```

```{r}
colnames(surveys)
combined_full <- inner_join(surveys, species, by = "species_id") %>%
  inner_join(plots, by = "plot_id") 
colnames(combined_full)
combined_full %>% 
  filter(plot_type == "Control"| plot_type == "Long-term Krat Exclosure") %>% 
  filter(taxa == "Rodent") %>% 
  select(year, genus, species,weight, plot_type) %>% 
  filter(!is.na(weight)) 

colnames(combined_full)
tail(combined_full, 100)
"Long-term Krat Exclosure" %in% combined_full$plot_type

```
###Solo in class excercise 

filter by some criteria 
1. Create a data frame with only data for the “species_id” “DO”, with the columns "year", "month", "day", "species_id", and "weight".
```{r}
combined_full %>%
  filter(species_id == "DO")  %>%
  select(year,month, day, species_id, weight) -> data1 
```
2. Create a data frame with only data for species IDs "PP" and "PB" and for years starting in 1995, with the columns "year", "species_id", and "hindfoot_length", with no missing values for "hindfoot_length".
```{r}
combined_full %>% 
  filter(species_id == "PP" | species_id == "PB") %>% 
  filter(year >= 1995) %>% 
  filter(!is.na(hindfoot_length)) %>% 
  select(year, species_id, hindfoot_length) -> data2
```

3.Create a data frame with the average "hindfoot_length" for each "species_id" in each "year" with no null values.
```{r}
combined_full %>% 
  filter(!is.na(hindfoot_length))%>% 
  group_by(species_id, year) %>% 
  summarise(mean_hindfoot_length = mean (hindfoot_length, na.rm = TRUE))

```
4.Create a data frame with the "year", "genus", "species", "weight" and "plot_type" for all cases where the "genus" is "Dipodomys".
```{r}
combined_full %>% 
  filter(genus=="Dipodomys") %>% 
  select(year, genus, species, weight, plot_type)
```
5. Make a scatter plot with "weight" on the x-axis and "hindfoot_length" on the y-axis. Use a log10() scale on the x-axis. Color the points by "species_id". Include good axis labels.
This plot guives us Hind foot length from all the data combined from surveys, plots, adn species. 
there is a different plot color for each species id and the x axis is scaled at a log10 scale for a better closer view (more centered) of the data 
```{r}
library("ggplot2")
ggplot(data = combined_full, mapping = aes(x = weight, y = hindfoot_length, color = species_id)) + 
  geom_point() +
  scale_x_log10() +
labs(x = "Weight(g)", y = "Hindfoot Length (mm)", title = " Combined Data (Surveys, Plots, and Species)")
```

6. 
Make a histogram of weights with a separate subplot for each "species_id". Do not include species with no weights. Set the "scales" argument to "free_y" so that the y-axes can vary. Include good axis labels.

facet_wrap(~ ) function to create a subplot for each species
used the `filter(!is.na ())` function for the removal of all the null weight values. I filter the data before plotting. used ggplot function to map the plots and specified that i wanted histograms with the gom_histogram function. Finally used the labs function to add labels to the 
```{r}
combined_full %>% 
 filter(!is.na(weight)) %>% 
ggplot( mapping = aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~species_id, scales = "free_y") +
labs(x = "Weight(g)", y = "Number of Individuals Within a Species", title = " Combined Data (Surveys, Plots, and Species)")
```
7. 
single = sign is an assignment `,` is and `|` is or 
```{r}
combined_full %>% 
  filter(species_id == "PP" | species_id =="PB" | species_id == "DM") %>%
  filter(plot_type == "Control"| plot_type == "Long-term Krat Exclosure"| plot_type == "Short-term Krat Exclosure") %>%
ggplot(.,mapping = aes(x = weight, fill = species_id) , color = species_id )+
geom_histogram( ) +
facet_wrap(~plot_type) +
  labs(x = "Weight(g)", y = "Number of Individuals Within the Species")+
  guides(fill=guide_legend(title="Species Id")) +
  ggtitle("Distribution Comparison based on Exposure")
```
Last solo assignment 
Develop a data manipulation pipeline for the Portal surveys table that produces a table of data for only the three Dipodomys species ("DM", "DO", "DS").

The species IDs should be presented as lower case, not upper case.
The table should contain information on the date, the species ID, the weight and hindfoot length.
The data should not include null values for either weight or hindfoot length.
The table should be sorted first by the species (so that each species is grouped together) and then by weight, with the largest weights at the top.

pipeline for the Portal surveys table used thr mutate function with the `columnname =tolower()` function to override the exsisting species_id column to a species_id colum that had all values in lowercase 
finally used the arrange function to sort the table by species so that the species are grouped by name and then used the desc function with the input weight to arrange weight in a descending manner from greatest to lowest instead of the ascending default 

```{r}
library(magrittr)
library(dplyr)
read.csv("../raw-data/surveys.csv") %>% 
  filter(species_id == "DM" | species_id =="DO" | species_id == "DS") %>%
  select(month, day, year, species_id, weight, hindfoot_length) %>%
  filter(!is.na(weight) & !is.na(hindfoot_length)) %>% 
  mutate(species_id = tolower(species_id)) %>%
  arrange(species_id, desc(weight))

```

```{r, child = 'Vectors_and_data_frames.Rmd'}

```


