---
title: "small-mammals"
output:
  pdf_document: default
  html_document: default
date: '2022-10-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
For all the data manipulation we are going to use the 'dplyr' package 

to install a package we use the function install_packages()
but to use the function inside a package we need to load them to our environment we do this with the function 'library()'
```{r}
library(dplyr)
```


#Read teh data table 

```{r cars}
#getwd()
surveys <- read.csv(file = "../raw-data/surveys.csv")
head(surveys)
#View(surveys)
plots <- read.csv(file = "../raw-data/plots.csv")
head(plots)
species <- read.csv(file = "../raw-data/species.csv")
head(species)
```

# Subst a data table by columns 

Using the dplyr function 'subset()':

```{r}
select(surveys, year, species_id, weight)
surveys_short <- select(surveys, year, species_id, weight )
head(surveys_short)
```
Other ways to subset a data table by columns: 

```{r}
columns2keep <- c("year", "species_id", "weight")
surveys[,columns2keep]
```
## Add columns to our data 

we can use the function `mutate()`

Add an extra columb that has weight divided by 10 

```{r}
colnames(surveys)
surveys$weight_decimal <- surveys$weight/10

colnames(surveys)
```

we can use function `mutate()`
```{r}
surveys <- mutate(surveys,weight_decimal_2 =weight/10)
```

```{r}
surveys
```

```{r}
arrange(surveys, weight)
```

to arrange in a descending order : 
```{r}
arrange(surveys, desc(weight))
```


##Filter Data 

to filter unwanted data or by some specific value in our data we use the function `filter()`
! meand 
```{r}
filter(surveys, species_id == "DS")
filter(surveys, species_id !="DS")
```

We can also filter by multiple conditions : 

```{r}
temp_table <- filter(surveys, species_id <="DS", year < 1998)
tail(temp_table, 100)
```

the logical operator `&` `|` mean "AND" and "OR"
```{r}
filter(surveys, species_id == "DS" & year >= 1998)
```

```{r}
filter(surveys, species_id == "DS" | year >= 1998)
```

lets filter for values that are NA we need the function `is.na()`
```{r}
is.na(surveys$weight)
surveys$weight
filter(surveys, is.na(weight))

```

```{r}
is.na(surveys$weight)
surveys$weight
filter(surveys, !is.na(weight))
```

#Exercise 1: Data manipulation

```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
```
```{r}
subset1 <- select(surveys, year, month, day, species_id)
```
```{r}

surveys_subset <- filter(surveys, !is.na(weight))
nrow(surveys_subset)

surveys_subset <- mutate(surveys_subset, weight_kg = weight/1000)
nrow(surveys_subset)

head(surveys_subset)
surveys_subset <-select(surveys_subset, year, species_id, weight_kg)
surveys_subset_filtered <- filter(surveys_subset, species_id == "SH")
nrow(surveys_subset_filtered)
```
## Pipes 

`%>%`- `magrittr` package required 
`|>`inbuilt in R very new so we need the updated version of r 
```{r}
library(dplyr)
mean(x = c(1, 3, 65, 38, 100)) ##sequential code 
c(1, 3, 65, 38, 100) %>% mean()


sqrt(mean(x = c(1, 3, 65, 38, 100 )))
c(1, 3, 65, 38, 100) %>% mean() %>% sqrt()
```
```{r}
surveys <- read.csv(file = "../raw-data/surveys.csv")
```
```{r}
surveys1 <- surveys %>%
  mutate(weight_kg = weight/1000 ) %>%
   select(year,species_id, weight_kg ) %>%
  filter(!is.na(weight_kg))
colnames(surveys1)
is.na(surveys1$weight_kg) %>% any() 

 
```


```{r}
colnames(surveys2)
nrow(surveys2)

surveys2 <- surveys %>%
   select(year, month, day, species_id, weight) %>%
  filter(species_id == "SH") %>%
  filter(!is.na(weight)) %>% 
  select(year, month, day, species_id) 

   

surveys2$species_id == "SH"

colnames(surveys2)
  
```

```{r}
ds_data <-  filter(surveys, species_id == "DS",!is.na(weight)) %>%
  ds_data_by_year <- arrange(ds_data, year) %>%
  ds_weight_by_year <- select(ds_data_by_year, year, weight)
```

```{r}
ds_data <- filter(surveys, species_id == "DS") %>% 
  filter(!is.na(weight)) 
ds_data_by_year <- arrange(ds_data, year)
ds_weight_by_year <- select(year, weight, species_id)
```

```{r}  
ds_weight_by_year <-surveys %>% 
  filter(species_id == "DS") %>% 
  filter(!is.na(weight)) %>% 
  arrange(year) %>% 
  select(year, weight, species_id)
nrow(ds_weight_by_year)
```

## Data Aggregation 
```{r}
surveys_by_year <- group_by(surveys, year, plot_id)
surveys_by_year
```
```{r}
plot_year_counts <- summarize(surveys_by_year, abundance = n())
```

```{r}  
plot_year_counts<-surveys %>% 
  group_by(plot_id, year) %>% 
  summarise(abundance = n())
```

Homework 
1
```{r}
surveys_individuals <- surveys %>%
  group_by(species_id) %>%
  summarize(abundance = n())

```
2
```{r}
surveys_individuals_year <- surveys %>%
  group_by(species_id, year ) %>%
  summarize(abundance = n())
```

3
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  filter(species_id == "DO") %>%
  group_by(year) %>%
  summarise(mean_year = mean(weight))
  

```















