---
title: "uhuru dataset"
output:
  pdf_document: default
  html_document: default
date: '2022-10-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 1. Describing the data that we are using 

We are using the data set from [this study](https://esapubs.org/archive/ecol/E095/064/metadata.php)

The Data From UHURU Acacia Survey is based on a study of assessing both direct and indirect consequences of the removal of nonrandom species of the environment. The UHURU consists of three wired electric fences to provoke herbivore exclusion treatments this also has a control group which is un-fenced. The units that we are using are for height meters and for weight kg. These three  Fenced treatments include "Mega" excludes elephants and giraffes only; "Meso" excludes both megaherbivores and mesoherbivores, ca. 40 kg and larger; and"Total" excludes all herbivores > 5 kg.




### 2. reading the data table into R 


First make sure we are in the correct working directory "getwd()"
"/Users/atziri/Bio 195-197/Data Science"

if it is not right set the working directory with 'setwd()'

how do i test where the computer is ?

```{r}
read.csv(file = "/Users/atziri/Bio 195-197/Data Science/raw-data/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```
```{r}
read.csv(file = "../raw-data/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```
```{r}
r_proj_wd <- "/Users/atziri/Bio 195-197/Data Science"
r_chunk_wd <- getwd()
r_proj_wd == r_chunk_wd
r_chunk_wd
```
```{r}
acacia <-read.csv(file = "../raw-data/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```
#The two dots represents what we need to do to go one up to the folder neded 

# 3. explore our data set 
'head()' gives us the first six rows 

```{r}
head(acacia)
summary(acacia)
colnames(acacia)
row.names(acacia)
class(acacia$SURVEY)
#the sapplu function all owes to apply a function to a list of objects 
#a data frame is a list of vectors of the same lenght 
sapply(acacia, class)


```

make sure that everything htat should be a numeric value is a number one way to check is the 'summary()' command 

another way is using the type function 


```{r}
typeof(acacia [,"HEIGHT"])


acacia$HEIGHT
```

we identified a column that has problematic data 
we need to fix this 

## Cleaning our raw data 
### 3.2 assign 'NA' lable to missing

We are going to read the data table again, but we are going to assign 'NA' to the "dead value "that we dont want in our 
arguments are always plain text 

```{r}
acacia <- read.csv(file = "/Users/atziri/Bio 195-197/Data Science/raw-data/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t", na.strings = "dead")
```
### 4. Visualize our data 

For this we are using the 'ggplot' package. 
lets install and load it 


```{r}
# install.packages("ggplot2")
library(ggplot2)
```

Now We are going to create our first plotting 
layer with the function 'ggplot. 

```{r}
colnames(acacia)
acacia$CIRC
```
The ggplot function creates  a blank canvas. the canvas contains our data 
and the variables that we plot. 

The aes function allows for the plugging of all lines colors and data we want to read 

A scatter plot can be created with the function (geom_point)
it can have a blank () however if you want to specidy color sixe and trasparency you add it
```{r}

```
```{r}
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT )) + geom_point(size = 1, color = "blue", alpha = 0.8, shape = 11) + 
  labs(x = "Circumference", y = "Height", title = "Data From UHURU Acacia Survey ")
```
to rescale th eplotting of the axis to log scale we use the function 'scale_y_log_10()'

```{r}
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT )) + 
  geom_point(size = 1, color = "purple", alpha = 0.4) + 
  scale_x_log10() +
  scale_y_log10()
```
```{r}
acacia$TREATMENT
```
```{r}
ggplot(data = acacia, mapping = aes (x = CIRC, y = HEIGHT, color = TREATMENT)) + 
 geom_point() + 
  facet_wrap(~TREATMENT)
```

### 4.2 Visualize a statystical analysis of correlation 

# Subplots or facets 

Teh function ro create subplots by a third variablr is called 'facet_wrap()'

```{r}
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT, color = TREATMENT)) + geom_point() + 
  facet_wrap(~TREATMENT)
```
The total that excludes all the herbivores has more trees than the fenced spaces with select herbivores 
meso has taller trees with larger circumference 
mega has shorter trees with smaller circumference 


A little interpretation of what is going on here 

How to test a Hypothesis 

## Model Filling functions 

The geom_smooth() function 

```{r}
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT, color = TREATMENT, linetype  = TREATMENT)) + geom_point() + 
 geom_smooth(methood = "loess") +
  facet_wrap(~TREATMENT)
```

## Histograms and barplots 

For bar plots use the geom_bar() function: 


```{r}
ggplot(data = acacia, aes(x = TREATMENT))+
  geom_bar()
```

If you want to see the distribution of a continuous variable we use the geom_ histogram() function:

```{r}
ggplot(data = acacia, mapping = aes(x = HEIGHT)) +
  geom_histogram()
```

The 'fill =' command gives the inside of the bar color and the 
'color = "" ' command gives the outline color of the bars  
```{r}
ggplot(data = acacia, mapping = aes(x = HEIGHT, fill = TREATMENT)) +
  geom_histogram(bins = 20, color = "black")
```



## layer multiple data from the same or different data sets 
```{r}
ggplot() +
  geom_point(data = acacia,
             mapping = aes(x = CIRC, y = HEIGHT, 
                           color = TREATMENT)) +
  geom_smooth(data = acacia, 
              mapping = aes(x = CIRC, y = HEIGHT))
```

# Save images as files 

```{r}
ggsave(filename = "acacia_by_treatment.pdf")
```



# In class exercise 1. 
# 1. 

```{r}
ggplot() +
  geom_point(data = acacia,
             mapping = aes(x = CIRC, y = AXIS1, )) +
  labs(x = "Circumference", y =  "Canopy Diameter", title = "Data From UHURU Acacia Survey ")
```
# 2. 
```{r}
ggplot() +
  geom_point(data = acacia,
             mapping = aes(x = CIRC, y = AXIS1, )) +
  labs(x = "Circumference", y =  "Canopy Diameter", title = "Data From UHURU Acacia Survey ") +
  scale_x_log10() +
  scale_y_log10()
```
# 3. 
```{r}
ggplot() +
  geom_point(data = acacia,
             mapping = aes(x = CIRC, y = AXIS1,
                           color = ANT)) +
  labs(x = "Circumference", y =  "Canopy Diameter", title = "Data From UHURU Acacia Survey ")
```
# 4. 
```{r}
ggplot() +
  geom_point(data = acacia,
             mapping = aes(x = CIRC, y = AXIS1,
                           color = ANT)) +
  labs(x = "Circumference", y =  "Canopy Diameter", title = "Data From UHURU Acacia Survey ") +
geom_point() + 
  facet_wrap(~ANT)
```
5. 
```{r}
ggplot(data = acacia,
             mapping = aes(x = CIRC, y = AXIS1,
                           color = ANT)) +
  labs(x = "Circumference", y =  "Canopy Diameter", title = "Data From UHURU Acacia Survey ") +
geom_point() +
  geom_smooth()+
  facet_wrap(~ANT) 
```


#Exercise 2. Histograms 

# 1. 

```{r}
ggplot(data = acacia, aes(x = ANT))+
  geom_bar()
```

# 2. 
```{r}
ggplot(data = acacia, mapping = aes(x = HEIGHT)) +
   labs(x = "Height (m)", y =  "Number of Acacia", title = "Data From UHURU Acacia Survey ") +
  geom_histogram(binwidth = .20 , color = "black")
```
# 3. 

```{r}
ggplot() +
  geom_histogram(data = acacia,
             mapping = aes(x = AXIS1) , alpha = 0.3, fill = "red") +
  geom_histogram(data = acacia, 
              mapping = aes(x = AXIS2), alpha = 0.3, fill = "black")
```
# 4. 
```{r}
ggplot() +
  geom_histogram(data = acacia,
             mapping = aes(x = AXIS1) , alpha = 0.3, fill = "red") +
  geom_histogram(data = acacia, 
              mapping = aes(x = AXIS2), alpha = 0.3, fill = "black") +
  facet_wrap(~ TREATMENT)
```


 Home exercises 

# Exercised 3. 

#1 To installed a package use the function 'install.packages("")'
```{r}
#install.packages("readr")

```
function library allows me to open the directory where the package i just downloaded is stored 
```{r}
library(readr)
```

getting and setting my working directory 

```{r}
setwd("~/Bio 195-197/Data Science/raw-data")
getwd()
```
```{r}
read_tsv(file = "/Users/atziri/Bio 195-197/Data Science/raw-data/TREE_SURVEYS.txt")
```
```{r}
read_tsv(file = "../raw-data/TREE_SURVEYS.txt")
```

#2 reading the file of the data set with function read_tsv and then assign it to the trees object 
```{r}
trees <- read_tsv(file = "../raw-data/TREE_SURVEYS.txt",
                  col_types = list(HEIGHT = col_double(),
                                AXIS_2 = col_double())) 
```

function gives me a summary of thw object trees which is now the dataset needed for the assignment 
```{r}
summary(trees)
```

#3
add a new column with the command 'nameofdataframehere$newcolnamehere' then give it the command for the value we need
the estimated canopy area calculated as the value in the AXIS_1 column times the value in the AXIS_2 column.
```{r}
trees$canopy_area <- trees$AXIS_1 * trees$AXIS_2
```

#4
Subset the trees data frame with just the SURVEY, YEAR, SITE, and canopy_area columns. 
```{r}
subset(trees, select = c('SURVEY' , 'YEAR' , 'SITE' , 'canopy_area'))
```

#5
creating a scatter plot using the ggplot() function. using the mapping function to add the aesthetics of the plot such as the data used for the x and y axis 
using the 'geom_point()'function to with the command size to add a size to the plots 
using the facet_wrap(~SPECIES ) function to create a subplot for each species. I used the function labs to add the labeling of each of the axis and the title of the graph. 
```{r}
ggplot(data = trees,
       mapping = aes(x = canopy_area, y = HEIGHT, color = TREATMENT )) +
  geom_point( size = 2) + 
  facet_wrap(~SPECIES ) +
 labs(x = "Canopy Area (m)", y = "Height (m)", title = "Data From Trees ")
```
#6 
logical vector that evaluates all rows that are in AXIS_1 and AXIS_2 that are below 20 
this wa to eliminate outliers all values above 20. 
```{r}
trees_2 <- trees$AXIS_1 <= 20 | trees$AXIS_2 <= 20 
summary(trees_2)
nrow(trees)
```

```{r}
trees <- trees[trees_2, ]
```

```{r}
ggplot(data = trees,
       mapping = aes(x = canopy_area, y = HEIGHT, color = TREATMENT )) +
  geom_point( size = 2) + 
  facet_wrap(~SPECIES ) +
 labs(x = "Canopy Area (m)", y = "Height (m)", title = "Data From Trees ")
```
#7 
used the pipe symbol %>% which takes th eoutput of the function and into another function as an argument this linked a sequence of analysis steps.  
Assignment needed to be used to the right rather than the left  
```{r}
library(dplyr)
trees %>%
group_by(YEAR, SPECIES) %>%
  summarize(species_abundance=n()) -> abundance 


```

#8 
Here i made a line plot with points by using the geom_line() and the  geom_point function 
used the facet_wrap(~SPECIES, scales = "free_y") command to see each trend clearly
```{r}
ggplot(data = abundance,
       mapping = aes(x = YEAR, y= species_abundance)) +
  geom_line()+
  geom_point()+ 
  facet_wrap(~SPECIES, scales = "free_y")

```
Exercise 4
Comparing  the circumference to height relationship in acacia and to the same relationship for trees in the region. This compares two data sets in a scatter plot 
I used the geom_points() function to specify the information we need from each of both data sets 
first are the comands of the trees data set- what we want on the back goes first 
the geom smooth command gives the smooth line to each one of the layers in the scatter plot 
the scale_x_log10() and  scale_y_log10() functions allow us to scale the data in the scatter plot 
we had the issue in office hrs to get the trend lines to be straight rather than curved. 
used the labs() functions to add labels to the scatter plot. 

```{r}
ggplot() +
  geom_point(data = trees, 
              mapping = aes(x = CIRC , y = HEIGHT), color = "gray") +
  geom_smooth(data = trees, 
              mapping = aes(x = CIRC , y = HEIGHT), color = "black") +
   scale_x_log10() +
  scale_y_log10()+
  geom_point(data = acacia,
             mapping = aes(x = CIRC , y = HEIGHT ), color= "red") +
  geom_smooth(data = acacia,
             mapping = aes(x = CIRC , y = HEIGHT ), color= "red") +
   scale_x_log10() +
  scale_y_log10()+
  labs(x = "Circumference (cm)", y = "Height (m)", title = "Data From Acacia VS Trees ")
  
```

